---
title: "Using database"
author: ""
date: "8/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Connect to the database

- Now let's use the databases
- First connect to the database

```{r}
library(DBI)
library(tidyverse)

db <- dbConnect(RSQLite::SQLite(), "~/covid.sqlite")
```

## Run queries

### Get sample rows from each table 

- Get sample rows from each table
- Use `dbGetQuery()`
  - Run SQL queries there

```{r}
dbGetQuery(db, "select * from covid limit 5")
dbGetQuery(db, "select * from demo limit 5")
dbGetQuery(db, "select * from trump limit 5")
```


### Bit more complex queries

1. Get the latest data for each county
2. Get the data on 1 Aug, 2020, and select county with more than 10 deaths

```{r}
dbGetQuery(db, "select * from covid where date = '2020-08-08'" )
dbGetQuery(db, "select * from covid where date = '2020-08-01' and deaths > 10" )

```
#### Ordering and aggregation

1. Order counties with the highest number of death
2. Calculate the state level death figures


```{r}
dbGetQuery(db, "select * from covid where date = '2020-08-08' order by deaths desc" )
dbGetQuery(db, "select state, sum(deaths) from covid where date = '2020-08-08' group by state" )
```


### Merging tables

- merge covid and demo table by fips
- (advanced) calculate per 1000 cases and deaths

```{r}
dbGetQuery(db, "select covid.*, demo.* from covid inner join demo on covid.fips = demo.fips")
dbGetQuery(db, "select covid.*, demo.*, covid.deaths / demo.totalPop as deaths_per1000  from covid inner join demo on covid.fips = demo.fips")
```

## dblyr

Now we will use `dbplyr`

### Get the object for each table

```{r}
tb_covid <- tbl(db, "covid")
tb_demo <- tbl(db, "demo")
```


### Let's check the same thing we did:


#### Get sample rows from each table 

- Get sample rows from each table
- Use `dbGetQuery()`
  - Run SQL queries there

```{r}
tb_covid %>% head(5)
tb_demo %>% head(5)
```


#### Bit more complex queries

1. Get the latest data for each county
2. Get the data on 1 Aug, 2020, and select county with more than 10 deaths

```{r}
tb_covid %>% filter(date == "2020-08-08")
tb_covid %>% filter(date == "2020-08-01") %>%
  filter(deaths > 10)

```
#### Ordering and aggregation

1. Order counties with the highest number of death
2. Calculate the state level death figures


```{r}
tb_covid %>%
  filter(date == '2020-08-08') %>% 
  arrange(desc(deaths))
tb_covid %>%
  filter(date == '2020-08-08') %>% 
  group_by(state) %>% summarize(tot_deaths = sum(deaths))
 
```


### Merging tables

- merge covid and demo table by fips
- calculate per 1000 cases and deaths


```{r}
tb_covid %>% left_join(tb_demo, by = "fips") %>%
  mutate(deaths_per1000 = deaths/TotalPop) 
```